---
title: "Linear Regression"
output: pdf_document
date: "2023-09-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

```{r}
Auto <- read_excel("C:/Users/I068117/UT_Machine Learning/Auto.xlsx")
names(Auto)
```

## Scatterplot of matrix of all variables

```{r }
# converting horsepower from character to numeric 
Auto$horsepower <- as.numeric(Auto$horsepower)
pairs(~ mpg + cylinders + displacement + weight + horsepower +
acceleration + year + origin, Auto)
```

## Correlation plot of all numeric variables

```{r }
# Selecting only the numeric X predictors

numeric_data <- Auto[sapply(Auto, is.numeric)]

# Remove rows with missing values
cleaned_data <- na.omit(numeric_data)

cor(cleaned_data)
```
## multiple linear regression with mpg as the response and all other variables except name as the predictors.

```{r }
mlr <- lm(mpg ~ cylinders + displacement + weight + horsepower +
acceleration + year + origin, data= cleaned_data )

summary (mlr)
```
## Interpretation

Q: Is there a relationship between the predictors and the response?
A: Yes

Q: Which predictors appear to have a statistically significant relationship to the response?
A: Predictors with less than 5 % p-value (p < 0.05) are statistically significant as we can reject the null hypothesis.

Displacement: The "displacement" predictor is statistically significant (p = 0.00844)

Weight: The "weight" predictor is  statistically significant (p < 0.0000000000000002)

Year: The "year" predictor is  statistically significant (p < 0.0000000000000002).

Origin: The "origin" predictor is  statistically significant (p = 0.000000467).


Q: What does the coefficient for the year variable suggest?
1. There is a positive relation between mpg & year
2. For 1 unit change in year, mpg increases by 0.750773

## Use the plot() function to produce diagnostic plots of the linear regression fit. Comment on any problems you see with the fit. Do the residual plots suggest any unusually large outliers?

```{r }
plot(mlr)
```
Interpretation:

The residual error is not normally distributed around the horizontal line and there is a curve/pattern in the distribution suggestion linear regression may not be a good fit. There are also large outliers in the residual plot

## Use the * and : symbols to fit linear regression models with interaction effects. Do any interactions appear to be statistically significant?

```{r }
#Interaction for all variables

interaction_model <- lm(mpg ~ cylinders * displacement * weight * horsepower *
acceleration * year * origin, data= cleaned_data )

# Print the summary of the model
summary(interaction_model)

```

Interpretation: No predictors seems statistically significant

```{r }
#  2 interactions of each variables

interaction_model2 =lm(mpg ~  (cylinders + displacement + weight + horsepower + 
    acceleration + year + origin) ^ 2 + cylinders:displacement:acceleration  ,data=cleaned_data )

# Print the summary of the model
summary(interaction_model2)

```
Interpretation: cylinders:displacement  , displacement:acceleration,horsepower:acceleration,horsepower:year ,acceleration:year cylinders:displacement:acceleration are statistically significant

```{r }
#  log (X)

interaction_model3 =lm(mpg ~  log(cylinders) + log(displacement) + log(weight) + 
    log(acceleration) + log(year) + log(origin),data= cleaned_data )

# Print the summary of the model
summary(interaction_model3)
```
Interpretation: log of weight, year, origin are statistically significant 

```{r }
#  3 interactions

interaction_model4 =lm(mpg ~  (cylinders + displacement + weight + horsepower + 
    acceleration + year + origin) ^ 3 ,data= cleaned_data )

# Print the summary of the model
summary(interaction_model4)

```
Few statistically significant interactions are: weight:year:origin  
horsepower:acceleration:year          
horsepower:acceleration:origin       
horsepower:year:origin     

```{r }
#  Polynomial for each variable to power of 2

interaction_model4 =lm(mpg ~  poly(cylinders, 2) + poly(displacement,2) + poly(weight,2) + 
    poly(acceleration,2) + poly(year,2) + poly(origin,2) ,data= cleaned_data )

# Print the summary of the model
summary(interaction_model4)

```
Interpretation: 2nd degree polynomial of weight, year, accelration are significant
```{r }
# SQRT

interaction_model5 =lm(mpg ~  sqrt(cylinders) + sqrt(displacement) + sqrt(weight) + 
    sqrt(acceleration) + sqrt(year) + sqrt(origin),data= cleaned_data )

# Print the summary of the model
summary(interaction_model5)

```

Interpretation: Sqrt of weight, year, origin are statistically significant